---
title: "P_xuthus"
author: "Khadijah Amarah Moosa"
date: "2025-02-26"
output: html_document
---

## Configuring knitr
I set up a R markdown document by configuring knitr options. I did this to ensure that my R code is visible in my final document but my setup chunk is hidden.

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load essential R libraries
I loaded the data from the CSV file and checked the first few rows and structure of the data. This was done to ensure the data was loaded correctly and to inspect its structure.

``` {r}
# Load the data (ensure the correct path or file is specified)
data <- read.csv("Papilio_xuthusRawData.csv")

# Check the first few rows and structure of the data
head(data)
str(data)
```


## Convert data to spacial objects
I converted the data to a spatial object by specifying the longitude and latitude columns. I then checked for any missing values in the longitude and latitude columns to ensure there were no missing coordinates.

``` {r}
# Assuming your CSV has 'lat' and 'long' columns
data_sf <- sf::st_as_sf(data, coords = c("longitude", "latitude"), crs = 4326)


sum(is.na(data_sf$longitude))  # Count missing longitude values
sum(is.na(data_sf$latitude))   # Count missing latitude values

```


## Check column names
I checked the column names of the spatial object to ensure the correct columns were included in the spatial data.

``` {r}
# Check the column names in your spatial object
colnames(data_sf)
```


## Create interactive map
I loaded the leaflet library and created an interactive map. I added tiles from a custom URL template and overlaid circle markers, which display the species information when clicked. I also added a key with the species name.

``` {r}
library(leaflet)

# Create an interactive map with customized basemap, blue markers, and adjusted legend with species name on the same line
leaflet(data_sf) %>%
  addTiles(urlTemplate = "https://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}.png") %>%  # Lighter basemap with more blue tones
  addCircleMarkers(
    popup = ~paste("<i>", species, "</i>"),  # Italicized species name in the popup
    color = "blue",  # Marker color kept as blue
    radius = 6,    # Marker size
    fillOpacity = 0.7
  ) %>%
  addLegend(
    position = "topright",  # Legend moved to the top right
    title = "Species", 
    colors = c("blue"), 
    labels = c("Papilio xuthus"),  # No HTML tags here to avoid the line break
    opacity = 1
  )

```

Figure 1. Interactive map displaying the locations of Papilio xuthus with species information shown in popups for each marker. The map uses a light basemap and geographic coordinates from the dataset.
